---
title: 'Pre-programme Assignment 1'
author: "Kostis"
date: "Mar 9, 2021"
output:
  html_document:
    highlight: zenburn
    theme: flatly
    toc: yes
    toc_float: yes
  word_document:
    toc: yes
---


```{r load-libraries, warning=FALSE, message=FALSE, echo=FALSE}
library(tidyverse)  # Load ggplot2, dplyr, and all the other tidyverse packages
library(gapminder)  # gapminder dataset
library(here)
```


# Task 1: Short biography written using markdown

You should write within this Rmd file a brief biography of yourself using markdown syntax. I know you have already achieved a lot, but a couple of paragraphs is more than enough.

To achieve full marks, you should include at least 4 of the following elements:

* Headers
* Emphasis (italics or bold)
* Lists
* Links
* Embeding images

> Please delete all the intro tet I wrote from line 22 to line 69 and start writing your short biography after this blockquote.

# Task 2: `gapminder` country comparison

You have seen the `gapminder` dataset that has data on life expectancy, population, and GDP per capita for 142 countries from 1952 to 2007. To get a glipmse of the dataframe, namely to see the variable names, variable types, etc., we use the `glimpse` function. We also want to have a look at the first 20 rows of data.

```{r}
glimpse(gapminder)

head(gapminder, 20) # look at the first 20 rows of the dataframe

```

Your task is to produce two graphs of how life expectancy has changed over the years for the `country` and the `continent` you come from. 

I have created the `country_data` and `continent_data` with the code below. 

```{r}
country_data <- gapminder %>% 
            filter(country == "Burkina Faso") # just choosing Greece, as this is where I come from

continent_data <- gapminder %>% 
            filter(continent == "Africa")
```

First, create a plot of life expectancy over time for the single country you chose. You should use  `geom_point()` to see the actual data points and `geom_smooth(se = FALSE)`  to plot the underlying trendlines. You need to remove the comments **#** from the lines below for your code to run.

```{r, lifeExp_one_country}
plot1 <- ggplot(data = country_data, mapping = aes(x = year, y = lifeExp))+
 geom_point() +
   geom_smooth(se = FALSE)+
   NULL 

 plot1
```

Next we need to add a title. Create a new plot, or extend plot1, using the `labs()` function to add an informative title to the plot.

```{r, lifeExp_one_country_with_label}
plot1 <- ggplot(data = country_data, mapping = aes(x = year, y = lifeExp))+
 geom_point() +
   geom_smooth(se = FALSE)+
   labs(title = "Life Expectancy in Burkina Faso",
        subtitle = "1952 - 2007",
       x = "Year",
       y = "Life Expectancy") +
   NULL


 print(plot1)
```


Secondly, produce a plot for all countries in the *continent* you come from. (Hint: map the `country` variable to the colour aesthetic).  

```{r lifeExp_one_continent}
 ggplot(data = continent_data , 
        mapping = aes(x = year , 
                      y = lifeExp, 
                      colour= country ))+
   geom_point() + 
   geom_smooth(se = FALSE) +
   NULL
```

Finally, using the original `gapminder` data, produce a life expectancy over time graph, grouped (or faceted) by continent. We will remove all legends, adding the `theme(legend.position="none")` in the end of our ggplot.

```{r lifeExp_facet_by_continent}
 ggplot(data = gapminder , mapping = aes(x = year , y = lifeExp  , colour=country ))+
   geom_point() + 
   geom_smooth(se = FALSE, alpha = 0.3) +
   facet_wrap(~continent) +
theme_minimal()+
   theme(legend.position="none") + #remove all legends
   NULL
```

Given these trends, what can you say about life expectancy since 1952? Again, don't just say what's happening in the graph. Tell some sort of story and speculate about the differences in the patterns.

> Type your answer after this blockquote.

# Task 3: Brexit voting

We will have a quick look at the results of the 2016 Brexit vote in the UK. First we read the data using `read_csv()` and have a quick glimpse at the data

```{r load_brexit_data, warning=FALSE, message=FALSE}
brexit_results <- read_csv(here::here("Data","brexit_results.csv"))


glimpse(brexit_results)
```


The data comes from [Elliott Morris](https://www.thecrosstab.com/), who cleaned it and made it available through his [DataCamp class on analysing election and polling data in R](https://www.datacamp.com/courses/analyzing-election-and-polling-data-in-r).

Our main outcome variable (or y) is `leave_share`, which is the percent of votes cast in favour of Brexit, or leaving the EU. Each row is a UK [parliament constituency](https://en.wikipedia.org/wiki/United_Kingdom_Parliament_constituencies). 

To get a sense of the spread of the data, plot a histogram and a density plot of the leave share in all constituencies.

```{r brexit_histogram, warning=FALSE, message=FALSE}
ggplot(brexit_results, aes(x = leave_share)) +
  geom_histogram(binwidth = 2.5)

ggplot(brexit_results, aes(x = leave_share)) +
  geom_density()


```

One common explanation for the Brexit outcome was fear of immigration and opposition to the EU's more open border policy. We can check the relationship (or correlation) between the proportion of native born residents (`born_in_uk`) in a constituency and its `leave_share`. To do this, let us get the correlation between the two variables

```{r brexit_immigration_correlation}
brexit_results %>% 
  select(leave_share, born_in_uk) %>% 
  cor()
```

The correlation is almost 0.5, which shows that the two variables are positively correlated. 

We can also create a scatterplot between these two variables using `geom_point`. We also add the best fit line, using  `geom_smooth(method = "lm")`. 

```{r brexit_immigration_plot}
ggplot(brexit_results, aes(x = born_in_uk, y = leave_share)) +
  geom_point(alpha=0.3) +
  geom_smooth(method = "lm") +
  theme_bw() +
  NULL
```

You have the code for the plots, I would like you to revisit all of them and use the `labs()` function to add an informative title, subtitle, and axes titles to all plots.

What can you say about the relationship shown above? Again, don't just say what's happening in the graph. Tell some sort of story and speculate about the differences in the patterns.

> Type your answer after, and outside, this blockquote.


# Submit the assignment

Knit the completed R Markdown file as ah HTML or Word document (use the "Knit" button at the top of the script editor window) and upload it to Canvas.


## Details

If you want to, please answer the following

- Who did you collaborate with: TYPE NAMES HERE
- Approximately how much time did you spend on this problem set: ANSWER HERE
- What, if anything, gave you the most trouble: ANSWER HERE
